Searching for long-period transiting planets in the \kepler\ light curves using supervised classification

### 2015-04-08 document/ms.tex
words for intro

\begin{itemize}

\item Description of the physical significance of these planets

\item Discussion of previous attempts (papers by Gaudi, Yee, Payne, Bakos,
\etc).

\item Comparison of method to filtering methods and an argument for why
supervised classification should be more robust to false alarms given a large
enough training set

\end{itemize}

@@

Assuming the stellar parameters provided by the NASA Exoplanet
Archive\footnote{\dfmtodo{URL for kepler stellar table; downloaded on SOME
DATE.}} \dfmtodo{cite Huber} and approximating the stellar noise using the
15-hour CDPP \dfmtodo{cite CDPP}, \dfmtodo{some figure} shows the extrapolated
number of transiting planets with $\period > 1500\,\mathrm{d}$ based on the
\citet{Dong:2013} power-law model.
For a detection threshold $f\sim10$, the expected number of single transit
events in the \kepler\ light curves is $\sim 150$.

>>>>

The transit method of exoplanet detection and characterization has been
demonstrated as a powerful method for building systematic catalogs of
exoplanets.
Despite the great success of the \kepler\ Mission with thousands of planet
discoveries \dfmtodo{cite}, current methods for exoplanet discoveries are
currently limited in the range of orbital periods that can be studied.
Specifically, the standard transit search procedures only discover signals
with at least three observed transits \dfmtodo{cite Jenkins, Petigura, \etc}.
For \kepler, with a baseline of about four years, this sets an upper limit on
the detectable periods of just over a year.
In the Solar System, Jupiter---with a period of 12 years---dominates the
planetary dynamics and, since it would only exhibit at most one transit in the
\kepler\ data, it would be missed by most existing transit search procedures.
It is possible to discover long-period planets like this using targeted radial
velocity (RV) surveys \dfmtodo{cite} but the cost of implementing a systematic
RV search is substantially higher than searching the existing and forthcoming
photometric data for single transits.

There are two main technical barriers to a search for single transit events.
The first is that the transit probability for long-period planets is very low;
scaling as $\propto\period^{-5/3}$ for orbital periods longer than the
baseline of contiguous observations.
Therefore, even if long-period planets are intrinsically common, they will
still be underrepresented in a transiting sample.
The second challenge is that there are substantial signals in the observed
light curves caused by stochastic processes---both instrumental
(pointing jitter, temperature variations, \etc) and astrophysical (stellar
variability, \etc)---that can masquerade as transit signals.
In practice, even using the most sophisticated systematics removal methods,
these false signal far outnumber the true single transits.

Nearly every transit search algorithm is built on the same principles and many
of the same decisions are made.
In particular, at the heart of most methods is a matched filtering step where
the likelihood of an approximate transit model is computed on a grid in the
physical parameters \dfmtodo{cite}.
Using these methods, and substantial hand-curation, some long-period
transiting candidates have been published \citep[for example][]{Batalha:2013,
Huang:2013, Kipping:2014a}.
For more recent data releases, the community has settled on more conservative
selection criteria where candidates are required to have multiple transits
\citep[for example][]{Petigura:2013, Burke:2014}.
This means that there has never been a systematic search for single transits
in the full \kepler\ dataset.

A qualitatively different approach to planet search is employed by the
\project{Planet Hunters} (PH)
project\footnote{\url{http://www.planethunters.org/}} \citep{Fischer:2012}.
PH is a ``citizen science'' project where visitors to the website look at
sections of light curve and mark the locations of transits that can be
identified visually.
Visual inspection can be a useful search technique for large single transits
because humans are able to robustly distinguish transit signals from the noise
and this project has yielded some promising long-period candidates and
confirmed planets \citep[for example][]{Wang:2013}.
One shortcoming of the PH method is that it can be difficult to fully quantify
the performance (completeness and reliability) of the search and PH does not
evaluate their users using synthetic transit signals as is now common practice
in the transit search literature.
This means that the PH sample of candidates cannot be used for robust
population inference.

We propose a transit search method that shares some similarities with the PH
model but, instead of human classifiers, we use a supervised classification
model trained on large numbers of simulated transit signals injected into
real \kepler\ light curves.
This procedure is motivated because, by definition, most randomly selected
sections of light curve contain no transits and there is an excellent physical
model for the signals of interest.
If we can train a classification model to robustly separate light curve
sections with transits from sections without then we should be able to use
this in-place of the humans to find single transit events.

@@

Assuming the stellar parameters provided by the NASA Exoplanet
Archive\footnote{\dfmtodo{URL for kepler stellar table; downloaded on SOME
DATE.}} \dfmtodo{cite Huber} and approximating the stellar noise using the
15-hour CDPP \dfmtodo{cite CDPP}, \dfmtodo{some figure} shows the extrapolated
number of transiting planets with $1500\,\mathrm{d} < \period <
5000\,\mathrm{d}$ based on the \citet{Dong:2013} power-law model.
For a detection threshold $f\sim10$, the expected number of single transit
events in the \kepler\ light curves is $\sim 60$.
